Creating the ``r params$package_name`` R package

### 2025-10-08 create-FSHybridPLS.Rmd
btwn_normalize_train_test

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

```{r}
#' --- Dependencies for Testing ---
library(fda) # Required for mock functional data

# Placeholder for a minimal predictor_hybrid constructor for testing purposes

# --- Test Script Start ---

# 1. Setup Mock Data
N_train <- 100
N_test <- 50
N_scalar <- 2
N_functional <- 1

# Create Functional Data (Needs to be present but remain untouched)
basis_obj <- create.bspline.basis(rangeval = c(0, 10), nbasis = 7)
train_coefs <- matrix(rnorm(N_train * 7), nrow = 7)
test_coefs <- matrix(rnorm(N_test * 7), nrow = 7)
mock_fd_train <- fd(train_coefs, basis_obj)
mock_fd_test <- fd(test_coefs, basis_obj)

# Create Scalar Data (Crucial for the test)
# Training data has a clear mean and SD
Z_train <- matrix(rnorm(N_train * N_scalar, mean = c(5, 10), sd = c(2, 3)), 
                  nrow = N_train, ncol = N_scalar, byrow = TRUE)
# Testing data has different properties (must be normalized by train stats)
Z_test <- matrix(rnorm(N_test * N_scalar, mean = c(1, 1), sd = c(10, 10)), 
                 nrow = N_test, ncol = N_scalar, byrow = TRUE)

# Create Hybrid Objects
W_train <- predictor_hybrid(Z = Z_train, functional_list = list(mock_fd_train))
W_test <- predictor_hybrid(Z = Z_test, functional_list = list(mock_fd_test))

# Store original Z for comparison
Z_train_original <- W_train$Z
Z_test_original <- W_test$Z

# 2. Run the Function Under Test
W_normalized_list <- scalar_normalize_train_test(W_train, W_test)
W_train_norm <- W_normalized_list$train
W_test_norm <- W_normalized_list$test

# 3. Verification and Assertions

cat("--- Running Verification Checks ---\n")

# Check 1: Training set is correctly normalized (Mean ~ 0, SD ~ 1)
train_mean_norm <- colMeans(W_train_norm$Z)
train_sd_norm <- apply(W_train_norm$Z, 2, sd)

stopifnot(
  "Train mean Col 1 not near 0" = all(abs(train_mean_norm[1]) < 1e-10),
  "Train mean Col 2 not near 0" = all(abs(train_mean_norm[2]) < 1e-10),
  "Train SD Col 1 not near 1" = all(abs(train_sd_norm[1] - 1) < 1e-10),
  "Train SD Col 2 not near 1" = all(abs(train_sd_norm[2] - 1) < 1e-10)
)
cat("Check 1: Training set Z-score normalization successful (Mean~0, SD~1).\n")

# Check 2: Functional components are untouched
stopifnot(
  "Functional list should not change" = isTRUE(all.equal(
    W_train$functional_list[[1]]$coefs, 
    W_train_norm$functional_list[[1]]$coefs
  ))
)
cat("Check 2: Functional components successfully preserved.\n")

# Check 3: Test set is normalized using TRAIN statistics (crucial consistency check)

# Recalculate the training statistics
train_mean_recalc <- get_mean(Z_train_original)
train_sd_recalc <- get_sd(Z_train_original)

# Manually normalize the test set using TRAIN stats
Z_test_expected <- t( (t(Z_test_original) - train_mean_recalc) / train_sd_recalc)

# Compare the function's output (W_test_norm$Z) against the manual calculation
stopifnot(
  "Test set normalization mismatch" = isTRUE(all.equal(W_test_norm$Z, Z_test_expected))
)
cat("Check 3: Test set correctly normalized using only Training statistics.\n")

cat("--- All tests passed successfully! ---\n")

@@

#
# scalar_normalize_train_test <- function(predictor_train, predictor_test){
#   predictor_train_normalized <- predictor_train
#   predictor_test_normalized <- predictor_test
#
#   mean_train <- get_mean(predictor_train@Z) # train mean function
#   sd_train <- get_sd(predictor_train@Z)
#
#   predictor_train_normalized@Z <-
#     scalar_normalize(
#       predictor_train@Z,
#       mean_train,
#       sd_train
#       )
#
#   predictor_test_normalized@Z <-
#     scalar_normalize(
#       predictor_test@Z,
#       mean_train,
#       sd_train
#     )
#
#   return(
#     list(
#       train = predictor_train_normalized,
#       test = predictor_test_normalized
#     )
#   )
# }
```

```{r}
#
# btwn_normalize_train_test <- function(train, test){
#   #"Scale the scalar predictors so that the variability between the functional and scalar predictors are comparable:")
#   Z_s <- train@Z
#   train_normalized <- train
#   test_normalized <- test
#
#   omega <- 0
#   for (k in 1:train@n_functional){ # for kth functional predictor
#     omega <- omega + sum(diag(inprod(
#       train@functional_list[[k]],
#       train@functional_list[[k]]
#       )))
#
#   }
#     omega <- omega / sum((Z_s)^2)
#
#     train_normalized@Z <- sqrt(omega) * (train@Z)
#     test_normalized@Z <- sqrt(omega) * (test@Z)
#
#     return(
#       list(
#         train = train_normalized,
#         test = test_normalized
#       ))}
```

>>>>

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

```{r}
#' Scale the scalar predictors so that the variability between the
#' functional and scalar predictors are comparable.
#'
#' This method calculates a scaling factor (sqrt(omega)) from the training
#' set, defined as the ratio of the total sum of squares (L2 norms) of all
#' functional components to the total sum of squares of all scalar components.
#'
#' @param train A predictor_hybrid object representing the training data.
#' @param test A predictor_hybrid object representing the test data.
#'
#' @return A list containing the normalized 'train' and 'test' predictor_hybrid objects.
#' @export
btwn_normalize_train_test <- function(train, test){

  # Type checks
  stopifnot(
    "'train' must be of class 'predictor_hybrid'" = inherits(train, "predictor_hybrid"),
    "'test' must be of class 'predictor_hybrid'" = inherits(test, "predictor_hybrid")
  )

  # Create copies for modification
  train_normalized <- train
  test_normalized <- test

  # Z_s for clarity in calculation
  Z_s <- train$Z

  # --- 1. Calculate the numerator of omega (Sum of L2 norms of all functional samples in train) ---
  omega_numerator <- 0

  for (k in seq_len(train$n_functional)){ # for kth functional predictor
    # fda::inprod(fd, fd) returns the N x N matrix of inner products.
    # The diagonal elements are the L2 norm squared for each sample.
    functional_inprod_matrix <- fda::inprod(
      train$functional_list[[k]],
      train$functional_list[[k]]
    )

    # Sum the diagonal to get the total squared norm for this predictor k
    omega_numerator <- omega_numerator + sum(diag(functional_inprod_matrix))
  }

  # --- 2. Calculate the denominator of omega (Sum of squares of all scalar samples in train) ---
  omega_denominator <- sum(Z_s^2)

  # Handle case where scalar variance is zero (prevent division by zero)
  if (omega_denominator == 0) {
    warning("Scalar component sum of squares is zero. No scaling applied.")
    omega <- 1
  } else {
    # Calculate the scaling factor omega
    omega <- omega_numerator / omega_denominator
  }

  # --- 3. Apply scaling factor sqrt(omega) to both train and test scalar components ---
  scaling_factor <- sqrt(omega)

  train_normalized$Z <- scaling_factor * (train$Z)
  test_normalized$Z <- scaling_factor * (test$Z)

  return(
    list(
      train = train_normalized,
      test = test_normalized
    )
  )
}

@@
